Geographical Python Teaching Resources: GeoPyTeR

### 2021-01-04 Geopyter.tex
implemented final revisions requested. Not sure that link to GitHub page fulfills the replicability request but it moves us closer.

University educators in many disciplines are playing catch-up, but Geography---thanks to its long relationship with computation \citep{ArribasBel2018}---has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago. At the same time, there is evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what used to be done with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline). However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present one means of bringing the benefits of (spatial) data science-approaches to instruction using Jupyter `notebooks' for the teaching of the Python programming language and geospatial analysis.

@@

The overarching purpose of this software is therefore to address two seemingly contradictory issues: the recognition that no two instructors teach in precisely the same way, and the fact that few instructors have the luxury of both time and tenure to develop compelling new course material in splendid isolation. This is  where the Geographical Python Teaching Resource (\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although developed with geographers in mind, it provides a generic means by which instructors in any discipline can selectively incorporate and re-mix programming and conceptual content from existing Jupyter notebooks while providing their own `gloss' for this content. Our approach therefore seeks to develop a rich community of developers and teachers who work \emph{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and who \emph{support} instructors in quickly adapting existing materials to new delivery formats.

@@

Without realising it, we were plugging into a larger framework of thinking emerging from the literature on Open Educational Resources (\textsc{oer}), which began as largely `legal and economic concept' designed to support royalty and license free access to educational resources \citep{butcher2010open}. It should be clear from this definition that there is no \emph{necessary} link to \textsc{foss} since a free e-book can qualify as an \textsc{oer}. However, these two areas are conceptually aligned in many ways, not least through their shared interest in the `4 Rs': reuse, redistribution, revision, and remixing \citep{Hilton2010}. Within \citeauthor{du2017}'s \citeyearpar{du2017} \textsc{oer} typology \gp is clearly `open courseware', but it is conceptually distinct from \textsc{mit}'s eponymous OpenCourseWare or Standford's Coursera since it is designed to promote what \citeauthor{Ehlers2011} calls `Open Educational Practice' \citeyearpar{Ehlers2011} in a manner not altogether dissimilar to that the Geographic Science \& Technology Body of Knowledge \citep{bok2018}.

@@

>>>>

University educators in many disciplines are playing catch-up, but Geography---thanks to its long relationship with computation \citep{ArribasBel2018}---has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago. In academia, there has been a `turn' towards coding (\eg \citealp{Brunsdon:2020aa}), but this has not been matched by changes in our teaching practice and there is evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what used to be done with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline). However, most countries have actually seen an increase in `geospatial jobs' \citep{Solis:2020aa} indicating unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques. In this article we present one means of bringing the benefits of (spatial) data science-approaches to instruction using Jupyter `notebooks' for the teaching of the Python programming language and geospatial analysis.

@@

The overarching purpose of this software is therefore to address two seemingly contradictory issues: the recognition that no two instructors teach in precisely the same way, and the fact that few instructors have the luxury of both time and tenure to develop compelling new course material in splendid isolation. This is  where the Geographical Python Teaching Resource (\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although developed with geographers in mind, it provides a generic means by which instructors in any discipline can selectively incorporate and re-mix programming and conceptual content from existing Jupyter notebooks while providing their own `gloss' for this content. Our approach therefore seeks to develop a rich community of developers and teachers who work \emph{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and who \emph{support} instructors in quickly adapting existing materials to new delivery formats. It should also be noted that our tool is fully compatible with notebooks written in other languages, such as R and its powerful geospatial analytics framework (see \citealp{Bivand:2020aa} for an overview).

@@

Without realising it, we were plugging into a larger framework of thinking emerging from the literature on Open Educational Resources (\textsc{oer}), which began as largely `legal and economic concept' designed to support royalty and license free access to educational resources \citep{butcher2010open}. It should be clear from this definition that there is no \emph{necessary} link to \textsc{foss}, or to open and reproducible research \citep{Brunsdon:2020aa}, since a free e-book can qualify as an \textsc{oer}. However, these two areas are conceptually aligned in many ways, not least through their shared interest in the `4 Rs': reuse, redistribution, revision, and remixing \citep{Hilton2010}. Within \citeauthor{du2017}'s \citeyearpar{du2017} \textsc{oer} typology \gp is clearly `open courseware', but it is conceptually distinct from \textsc{mit}'s eponymous OpenCourseWare or Standford's Coursera since it is designed to promote what \citeauthor{Ehlers2011} calls `Open Educational Practice' \citeyearpar{Ehlers2011} in a manner not altogether dissimilar to that the Geographic Science \& Technology Body of Knowledge \citep{bok2018}.

@@

\section{Supplemental}

Demonstration notebooks can be found in the `sessions' directory of the \gp project on GitHub: \href{https://github.com/pysal/geopyter/tree/master/sessions}{github.com/pysal/geopyter/tree/master/sessions}.
